\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-07-25 forwards-and-backwards.tex
adjustments to main text

\comment{TODO: talk about $N \to \infty$ rather than $\theta \to \infty$.}

@@

\begin{remark}
Although this model allows fairly general birth and death mechanisms,
there are a number of limitations.
Perhaps most obviously, to simplify the notation
individuals give birth to only one offspring at a time,
although this restriction could be easily lifted
\citep[as in][]{etheridge/kurtz:2018}.
Furthermore, individuals do not move during their lifetime,
and the age of an individual does not affect their fecundity or death rates.
Finally, there is no notion of mating,
so availability of mates will not restrict the spread of a population,
and the lineages we follow will be uniparental.
For these reasons, the model is most obviously applicable to bacterial populations or selfing plants,
although we do not anticipate that incorporation of these complications
will change the general picture.
\end{remark}

@@

\begin{definition}[Model parameterization and assumptions]
\label{def:model_setup}
Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
are uniformly continuous functions
from $\IR^d \times \IR_{\ge 0}$ to $\IR$.
Furthermore, suppose $r$ takes values in $[0, 1]$
and has uniformly bounded second derivatives,
that $\gamma(x,m) \ge 0$,
and that $m^2 \gamma(x, m)$ is uniformly bounded.
Fix positive values $\theta$, $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
with $\epsilon_r < \epsilon_\gamma$.
Suppose $F(x,m)$ is bounded above
with uniformly bounded first and second derivatives in $m$,
and define
\begin{align} \label{eqn:mu_defn}
    \mu(x, \eta)
    =
    r(x, \smooth{r} \eta(x))
    \gamma(\smooth{\gamma} \eta(x))
    - \frac{1}{\theta}
    F(x, \smooth{F} \eta(x)) .
\end{align}
Although the models will make sense without this requirement,
for the purposes of convergence we will also require that $\theta > 1/(\epsilon_\gamma - \epsilon_r)$.
\end{definition}

@@

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction},
For each value of $N$ and $\theta = \theta(N)$,
we define $(\eta^N_t)_{t \geq 0}$ to be the solution to the Martingale Problem,
with initial condition $\eta^N_0 \in \measures$,
such that for all smooth functions $f \in C^{\infty}_{0}(\mathbb{R}^d)$,
writing $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$
and $q_\theta(x, dy)$ as the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eq: Pre-limit Martingale Characterisation}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
\\ & \qquad \qquad \qquad \qquad {}
        \times \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale with quadratic variation
\begin{equation}
\begin{aligned} \relax
[ M^N_t(f)] =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \langle \gamma(x, \eta^N_{s})
        \int f^2(y)r(y, \eta^N_{s}) q_\theta(x,dy) 
    , \eta^N_{s}(dx) \rangle \\
& \qquad \qquad {}
    + \langle \mu(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition \ref{def:model_setup}
    and assume that as $N \to \infty$, also $\theta \to \infty$
    in such a way that $\theta/N \to \alpha$.
    (However, interaction distances $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$
    remain constant.)
    The sequence of stochastic processes
    $(\eta^N_t)_{t \geq 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\eta_t)_{t \geq 0}$
    which satisfies that, for every $f \in C^\infty_0(\IR^d)$,
    \begin{align} \label{eqn:mgale_problem}
        \begin{split}
        M_t
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                \\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds,
        \end{split}
    \end{align}
    is a martingale.
    The quadratic variation of $M_t$, denoted $\langle M \rangle_t$, is given by 
    \begin{align} \label{eqn:mgale_variation}
        \langle M \rangle_t
        =
        \alpha
        \int_0^t
        \big\langle
            \left(\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right) + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    In other words, the limit is a measure-valued process with non-local interactions.
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

Recall when interpreting~\eqref{eqn:mgale_problem}
that for instance $r(x, \eta_s) = r(x, \smooth{r} \eta_s(x))$,
and so $\DG(fr)(x) = \DG(f(\cdot) r(\cdot, \smooth{r} \eta_s(\cdot)))(x)$.

@@

\begin{corollary}[Superprocess limits] \label{cor:superprocess_uniqueness}
    In the context of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations \eqref{eqn:mgale_problem} and \eqref{eqn:mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges to that solution
    as $N \to \infty$.
    \comment{
        Is this what we want to say here? Could be said better.
        TODO: see what Dawson \& Li say about uniqueness
        so we have at least some conditions under which this happens.
    }
\end{corollary}

@@

\comment{TODO: find some conditions under which we have convergence to a unique solution?}
\comment{TODO: Tom has conditions (coefficient are Lipschitz?) under which
    the solution to \eqref{eqn:nonlocal_pdf} has a density (Thm 2.1 in DDMOT);
    should we insert those here (as a Corollary or Proposition, maybe?)}

@@

Now we will present a lookdown construction for the general process
(i.e., of Definition \ref{defn:mgale_construction}),
in the spirit of~\cite{kurtz/rodrigues:2011}. 
The general set-up is as follows.
Each individual will be labeled with a ``level'',
which will be a number in $[0, N]$.
We will still encode the process embellished by these levels
as a point measure:
if the $i^\mathrm{th}$ individual's spatial location is $x_i$
and level is $u_i$, then we will write:
$$
    \lp = \sum_i \delta_{x_i, u_i} ,
$$
which is a measure on $\IR^d \times [0, N]$.
Note that each individual contributes mass 1 to the measure,
not $1/N$ as above.
In some sense, these levels do not affect the process:
the pushforward of this measure-valued ``spatial-level process'',
divided by $N$,
will agree with the purely spatial process defined above.
Furthermore, the levels of the individuals present at any spatial location
will be conditionally uniform on $[0, N]$.
However, an individual's level in some sense encodes their future reproductive output:
individuals with lower levels tend to live longer, and have more offspring.
For more explanation of the set-up and how this is possible,
see \citet{etheridge/kurtz:2018} and \citet{kurtz/rodrigues:2011}
(and note that our $N$ corresponds to the $\lambda$ of those papers).
Furthermore, the limiting spatial-level process will still be a point measure,
and so we explicitly retain the notion of individuals and lineages through to the 
large-population-density limit.

@@

In this section,
we'll define the process $(\lp_t)_{t \ge 0}$ in terms of the dynamics of labeled particles,
and write down its generator.
Since the dynamics depend on the spatial locations of particles,
in this section we define $\eta_t$ to be the corresponding spatial measure,
i.e.,
$$
    \eta_t(\cdot) = \frac{1}{N} \lp_t(\cdot \times [0, N])  .
$$
A nontrivial consequence of the way we define $\lp_t$ will be that
this process has the same distribution as the process $(\eta_t)_{t \ge 0}$ defined above
in Definition~\ref{defn:mgale_construction}.
(This provides our justification for using the same notation for both.)
Finally, note that we first describe the dynamics of individuals
in the original time units -- i.e., before scaling by $\theta$,
which will appear below.

Suppose that the initial population is composed of $O(N)$ many particles
with levels uniformly distributed between $[0, N]$.
Suppose the current state of the population is $\lp$,
with spatial projection $\eta$.
An individual at spatial location $x$ with level $u$
produces one juvenile offspring at rate 
$$
2 \left(1 - \frac{u}{N}\right) \gamma(x, \eta) ,
$$
which disperses to a location relative to $x$ drawn from the kernel $q(x, \cdot)$.
(Averaging over the uniform distribution of the level $u$,
we recover the birth rate $\gamma(x, \eta)$.)
This juvenile -- suppose it's location is $y$ --
either survives, with probability $r(y, \eta)$, or immediately dies.
(As before, ``maturity'' is instantaneous.)

@@

The remaining term in the generator is due to the motion of particles' levels:
\begin{align} \label{eqn:level_generator}
    f(\lp)
    \mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    \left(
    \frac{\theta}{N}
        \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        b(x, \eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align}

@@

% % % % % % % %
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:lookdown_mgale}
For given positive values of $N$ and $\theta$,
we define $(\lp^{N}_t)_{t \geq 0}$
to be the unique solution
to the Martingale Problem
with initial condition $\lp_0 \in \mathcal{M}_F(\mathbb{R}^d \times [0,N])$
and generator $A^{N}$ that satisfies
\begin{equation*}
\begin{split}
& A^{N}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac{1}{2N} \int_u^N g(x,u_1) du_1
            \frac{
                \theta \int_{\mathbb{R}^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        \frac{\theta}{N} \int_u^N
        \int_{\mathbb{R}^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        \frac{\theta}{N} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) u^2 -b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation*}
for all smooth functions $g \in C^{2}_{0}(\mathbb{R}^d)$,
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$ is the scaled spatial pushforward of $\lp$,
as before.
\end{definition}

@@

\begin{theorem}
    \label{thm:lookdown_convergence}
    Let $(\lp_t^N)$ be as defined in Definition~\ref{defn:lookdown_mgale}
    and assume that as $N \to \infty$,
    $\theta \to \infty$ in such a way that $\theta/N \to \alpha$.
    The sequence of stochastic processes $(\lp_t^N)_{t \ge 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\lp_t)_{t \ge 0}$ which satisfies,
    for every $f: \IR^d \to \IR$
    \plr{TODO: write down limiting martingale characterization}
\end{theorem}

\comment{I don't think this bit quite makes sense yet without understanding the MMT
section (but, it should be understandable without understanding that section).}

@@

Although in Section \ref{sec:proofs} we will show convergence of the lookdown process
defined by Definition~\ref{defn:lookdown_mgale}
to a stochastic process in which individuals retain their identities,
it is convenient to construct the lookdown process
in a way that explicitly represents the motion of each individual and their offspring.
This will then make obvious how lineages move in the limit,
and will allow us to identify all 
individuals in the current population that are descendants of a
given ancestor at time zero. In theory at least, this allows us to
recover all the information about genealogies relating individuals 
sampled from the present day population. This idea draws on the notion
of `tracers', popular in statistical physics and used in population
genetics by a number of authors including 
\cite{biswas/etheridge/klimek:2018, durrett/fan:2016, hallatschek/nelson:2008}.

@@

@@

Potential reproduction events occur at times $\tau_a + s$
for each point $(s, v, w, \kappa) \in \Pi_a$.
We say `potential' since if the resulting level is greater than $N$,
the event does not happen.
If this is the $j^\text{th}$ point in $\Pi_a$,
the new label is $a \concat j$, the (potential) birth time is $\tau_{a \concat j} = \tau_a + s$.
Similarly, the potential birth location is 
$$
    y
    :=
    \frac{1}{\theta}\meanq(X_a(\tau_{a \concat j}-))
    +
    \frac{1}{\sqrt{\theta}}K(X_a(\tau_{a \concat j}-))z,
$$
where $K(x)K^{T}(x) = \covq(x)$.

>>>>

@@

\begin{remark}
Although this model allows fairly general birth and death mechanisms,
there are a number of limitations.
Perhaps most obviously, to simplify the notation
individuals give birth to only one offspring at a time,
although this restriction could be easily lifted
\citep[as in Section 3.4 of][]{etheridge/kurtz:2018}.
Furthermore, individuals do not move during their lifetime,
and the age of an individual does not affect their fecundity or death rates.
Finally, there is no notion of mating
(although limitations on reproduction due to availability of mates can be incorporated into the birth rate, $\gamma$),
so the lineages we follow will be uniparental.
For these reasons, the model is most obviously applicable to bacterial populations or selfing plants,
although we do not anticipate that incorporation of these complications
will change the general picture.
\end{remark}

@@

\begin{definition}[Model parameterization and assumptions]
\label{def:model_setup}
Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
are uniformly continuous functions
from $\IR^d \times \IR_{\ge 0}$ to $\IR$.
Furthermore, suppose $r$ takes values in $[0, 1]$
and has uniformly bounded second derivatives,
that $\gamma(x,m) \ge 0$,
and that $m^2 \gamma(x, m)$ is uniformly bounded.
Fix positive values $\theta$, $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
with $\epsilon_r < \epsilon_\gamma$.
Suppose $F(x,m)$ is bounded above
with uniformly bounded first and second derivatives in $m$,
and define
\begin{align} \label{eqn:mu_defn}
    \mu(x, \eta)
    =
    \max\left(0, 
        r(x, \smooth{r} \eta(x))
        \gamma(\smooth{\gamma} \eta(x))
        - \frac{1}{\theta}
        F(x, \smooth{F} \eta(x)) 
    \right) .
\end{align}
Although the models will make sense without this requirement,
for the purposes of convergence we will also require that $\theta > 1/(\epsilon_\gamma - \epsilon_r)$.
\end{definition}

@@

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction},
For each value of $N$ and $\theta = \theta(N)$,
we define $(\eta^N_t)_{t \geq 0}$ to be the solution to the Martingale Problem,
with initial condition $\eta^N_0 \in \measures$,
such that for all smooth functions $f \in C^{\infty}_{0}(\mathbb{R}^d)$,
writing $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$
and $q_\theta(x, dy)$ as the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eq: Pre-limit Martingale Characterisation}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
\\ & \qquad \qquad \qquad \qquad {}
        \times \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N_t(f)] =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \langle \gamma(x, \eta^N_{s})
        \int f^2(y)r(y, \eta^N_{s}) q_\theta(x,dy) 
    , \eta^N_{s}(dx) \rangle \\
& \qquad \qquad {}
    + \langle \mu(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition \ref{def:model_setup}
    and assume that as $N \to \infty$, also $\theta \to \infty$
    in such a way that $\theta/N \to \alpha$.
    (However, interaction distances $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$
    remain constant.)
    The sequence of stochastic processes
    $(\eta^N_t)_{t \geq 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\eta_t)_{t \geq 0}$
    which satisfies that, for every $f \in C^\infty_0(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                \\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds,
        \end{split}
    \end{align}
    is a martingale,
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            \left(\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right) + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    In other words, the limit is a measure-valued process with non-local interactions.
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

Recall when interpreting~\eqref{eqn:limiting_mgale_problem}
that for instance $r(x, \eta_s) = r(x, \smooth{r} \eta_s(x))$,
and so $\DG(fr)(x) = \DG(f(\cdot) r(\cdot, \smooth{r} \eta_s(\cdot)))(x)$.

@@

\begin{corollary}[Superprocess limits] \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations \eqref{eqn:limiting_mgale_problem} and \eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges to that solution
    as $N \to \infty$.
    \comment{
        Is this what we want to say here? Could be said better.
        TODO: see what Dawson \& Li say about uniqueness
        so we have at least some conditions under which this happens.
    }
\end{corollary}

@@

\comment{TODO: Tom has conditions, referring to \citet{KURTZ1999103}, (coefficient are Lipschitz?) under which
    the solution to \eqref{eqn:nonlocal_pdf} is unique and has a density (Thm 2.1 in DDMOT);
    insert those here (as a Corollary or Proposition)}

@@

Now we will present a lookdown construction for the general process
(i.e., of Definition \ref{defn:mgale_construction}),
in the spirit of~\cite{kurtz/rodrigues:2011}. 
The general set-up is as follows.
Each individual will be labeled with a ``level'',
which will be a number in $[0, N]$.
We will still encode the process embellished by these levels
as a point measure:
if the $i^\mathrm{th}$ individual's spatial location is $x_i$
and level is $u_i$, then we will write:
$$
    \lp = \sum_i \delta_{x_i, u_i} ,
$$
which is a measure on $\IR^d \times [0, N]$.
Note that each individual contributes mass 1 to the measure,
not $1/N$ as above.
In some sense, these levels do not affect the process:
the pushforward of this measure-valued ``spatial-level process'',
divided by $N$,
will agree with the purely spatial process defined above.
Furthermore, at any time, the levels of the individuals present at any spatial location
will be exchangeable and conditionally uniform on $[0, N]$:
in particular, choosing the $k$ individuals with the lowest level in some region of space
at some time is equivalent to taking a uniform random draw of size $k$ from the individuals in that region
for the purely spatial process.
However, an individual's level in some sense encodes their future reproductive output:
individuals with lower levels tend to live longer, and have more offspring.
For more explanation of the set-up and how this is possible,
see \citet{etheridge/kurtz:2018} and \citet{kurtz/rodrigues:2011}
(and note that our $N$ corresponds to the $\lambda$ of those papers).
Furthermore, the limiting spatial-level process will still be a point measure,
and so we explicitly retain the notion of individuals and lineages through to the 
large-population-density limit.

@@

In this section,
we'll define the process $(\lp_t)_{t \ge 0}$ in terms of the dynamics of labeled particles,
and write down its generator.
Since the dynamics depend on the spatial locations of particles,
in this section we define $\eta_t$ to be the corresponding spatial measure,
i.e.,
$$
    \eta_t(\cdot) = \frac{1}{N} \lp_t(\cdot \times [0, N])  .
$$
A nontrivial consequence of the way we define $\lp_t$ will be that
this process has the same distribution as the process $(\eta_t)_{t \ge 0}$ defined above
in Definition~\ref{defn:mgale_construction}.
(This provides our justification for using the same notation for both.)

As before, we first describe the dynamics of individuals
in the original time units -- i.e., before scaling by $\theta$,
which will be introduced below.
Suppose that the initial population is composed of $O(N)$ many particles
with levels uniformly distributed between $[0, N]$.
Suppose the current state of the population is $\lp$,
with spatial projection $\eta$.
An individual at spatial location $x$ with level $u$
produces one juvenile offspring at rate 
$$
2 \left(1 - \frac{u}{N}\right) \gamma(x, \eta) ,
$$
which disperses to a location relative to $x$ drawn from the kernel $q(x, \cdot)$.
(Averaging over the uniform distribution of the level $u$,
we recover the birth rate $\gamma(x, \eta)$.)
This juvenile -- suppose it's location is $y$ --
either survives, with probability $r(y, \eta)$, or immediately dies.
(As before, ``maturity'' is instantaneous.)

@@

The remaining term in the generator is due to the motion of particles' levels,
mapping $f(\xi)$ to
\begin{align} \label{eqn:level_generator}
    f(\lp)
    \sum_{(x, u) \in \lp}
    \left(
    \frac{\theta}{N}
        \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        b(x, \eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align}

@@

% % % % % % % %
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:lookdown_mgale}
For given positive values of $N$ and $\theta$
and $\lp_0 \in \mathcal{M}_F(\mathbb{R}^d \times [0,N])$,
we define $(\lp^{N}_t)_{t \geq 0}$
to be the unique solution to the martingale problem
with initial condition $\lp_0$ and generator $A^{N}$ that satisfies
\begin{equation*}
\begin{split}
& A^{N}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac{1}{2N} \int_u^N g(x,u_1) du_1
            \frac{
                \theta \int_{\mathbb{R}^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        \frac{\theta}{N} \int_u^N
        \int_{\mathbb{R}^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        \frac{\theta}{N} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) u^2 -b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation*}
for all smooth functions $g \in C^{2}_{0}(\mathbb{R}^d)$,
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$ is the scaled spatial pushforward of $\lp$,
as before.
\end{definition}

Since everything is finite,
it is clear that the martingale problem for finite $N$ has a unique solution.
Next we state the limiting martingale problem,
for which we do not have uniqueness (although we construct an explicit solution later).

\begin{definition}[Martingale Problem Characterisation, $N=\infty$]
    \label{defn:limiting_lookdown_mgale}
    For a given value of $\alpha \in [0, \infty)$,
    define the operator $A$ acting on test functions of the form
    $f(\xi) = \prod_{(x, u) \in \xi} g(x, u)$ with $g \in C_0^2(\IR^d \times [0, \infty)$
    and for which there exists a $u_0$ with $g(x, u) = 1$ for all $u > u_0$.
    Then, let $A f$ be
    \begin{align} \begin{split} \label{eqn:limiting_lookdown_generator}
    A f(\lp)
    &=
        f(\lp)  \sum_{(x, u) \in \xi}
        \gamma(x, \eta)
            \frac{
                \DG(g(\cdot, u) r(\cdot, \eta))(x) - g(x,u) \DG r(x,\eta)
            }{
                g(x, u)
            }
    \\ &\qquad {} +
        f(\lp) \sum_{(x, u) \in \xi}
        2 \alpha \gamma(x, \eta) r(x, \eta \int_u^\infty (g(x, u_1) - 1) du_1
    \\ &\qquad {} +
        f(\lp) \sum_{(x, u) \in \xi}
        \left(
            \alpha \gamma(x, \eta) r(x, \eta) u^2
            -
            \left\{
                \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta)
            \right\} u
        \right)
        \frac{\partial_u g(x, u)}{ g(x,u) }  .
    \end{split} \end{align}
    We say that a $\lpmeasures$-valued process $(\xi_t)_{t \ge 0}$
    is a solution to the $(A, \lp_0)$ martingale problem
    if it has initial distribution $\lp_0$
    and $f(\lp_t) - \int_0^t Af(\lp_s) ds$ is a martingale for all $f$
    as defined above.
\end{definition}

\comment{We talked about having the next paragraph be a Proposition or something?}

The lookdown processes have been carefully constructed so that
if we average over levels in the expression for the lookdown generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we get the generator for the population process
(visible in equation~\eqref{eq: Pre-limit Martingale Characterisation}
or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (and some boundedness conditions)
in Appendix \ref{sec: Markov Mapping Theorem Application},
the Markov Mapping Theorem \citep[Thm A.2 in][]{etheridge/kurtz:2018}
tells us that for any population process $(\eta^N_t)_{t \ge 0}$
(i.e., solution to the martingale problem of Definition~\ref{defn:mgale_construction})
% of equations~\eqref{eq: Pre-limit Martingale Characterisation} and \eqref{eqn:prelimit_martingale_variation}
there is a lookdown representation $(\lp^N_t)_{t \ge 0}$
(i.e., solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])$
and such that at each $t$, conditioned on $(\eta^N_s){s \le t}$, the levels of $\lp_t^N$
are independently and uniformly distributed on $[0, N]$.
The same property holds in the limit:
for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of Theorem~\ref{nonlocal_convergence}
there is a lookdown process $(\lp^N_t)_{t \ge 0}$
that is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
and such that at each $t$,
$\lp_t$ is conditionally Poisson with Cox measure $\eta_t(dx) \times du$,
conditioned on $(\eta_s)_{s \le t}$.
Furthermore, convergence of the lookdown processes would imply convergence of the population processes
\citep[Thm A.9 in][]{KR}.

@@

\begin{theorem}
    \label{thm:lookdown_convergence}
    Let $(\lp_t^N)$ be as defined in Definition~\ref{defn:lookdown_mgale}
    and assume that as $N \to \infty$,
    $\theta \to \infty$ in such a way that $\theta/N \to \alpha$.
    Suppose also that $\eta^N_0 \to \eta_0$ (as measures on $\IR^d$),
    and that for each $N$, $\lp^N_0(dx, du)$ is conditionally Poisson on $\IR^d \times [0, N]$
    with Cox measure $\eta^N_0(dx) \times du$.
    Then, the sequence of stochastic processes $(\lp_t^N)_{t \ge 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\lp_t)_{t \ge 0}$
    with $\lp_0$ conditionally Poisson with Cox measure $\eta_0(dx) \times du$,
    that is a solution to the martingale problem~\ref{defn:limiting_lookdown_mgale}.
\end{theorem}

\comment{TODO: remove this next bit?}

@@

In this section we explicitly construct the lookdown process
in a way that explicitly represents the motion of each individual and their offspring.
This provides an explicit construction of the solution to
the limiting martingale problem~\ref{defn:limiting_lookdown_mgale}.
It will furthermore make it obvious how lineages move in the limit,
and will allow us to identify all 
individuals in the current population that are descendants of a
given ancestor at time zero. In theory at least, this allows us to
recover all the information about genealogies relating individuals 
sampled from the present day population. This idea draws on the notion
of `tracers', popular in statistical physics and used in population
genetics by a number of authors including 
\cite{biswas/etheridge/klimek:2018, durrett/fan:2016, hallatschek/nelson:2008}.

@@

\comment{Perhaps the following is confusing having $t$ for the point of $\Pi_a$,
and we should instead use... $\tau$?}

@@

Potential reproduction events occur at times $t$
for each point $(t, v, w, \kappa) \in \Pi_a$ with $t \ge \tau_a$.
We say `potential' since if the resulting level is greater than $N$,
the event does not happen.
If this is the $j^\text{th}$ point in $\Pi_a$,
the new label is $a \concat j$, the (potential) birth time is $\tau_{a \concat j} = t$.
Similarly, the potential birth location is $y(X(t-), z)$, where
$$
    y(x, z)
    :=
    \frac{1}{\theta}\meanq(x)
    +
    \frac{1}{\sqrt{\theta}}K(x) z,
$$
where $K(x)K^{T}(x) = \covq(x)$.

Next we must choose the level of the offspring.
We would like the individual to produce offspring at $y$ at rate
$2 (1 - u/N) \theta \gamma(x, \eta) r(x + y, \eta)$,
and to do this
we will associate the point $(s, v, w, \kappa) \in \Pi_a$ with level $u + v \ell$,
where $\ell$ is chosen so that this is the rate of appearance of points in $\Pi_a$
with level below $N$.
Since the mean measure of $\Pi_a$ is Lebesgue in the $t$ and $v$ directions,
this requires that $v < (N - u) / \ell$, 
and so such points appear at rate $(N-u)/\ell$.
Therefore, we define
$$
    \ell(x, y, \eta)
    =
    \frac{
        1
    }{
        2 N^{-1} \theta \gamma(x, \eta) r(x + y, \eta)
    } ,
$$
and using this, the (potential) new level is
\begin{equation*}
    U_{a \concat j}(t)
    =
    U_a(t)
    +
    v \ell(X_a(t-), y(X_a(t-), z), \eta(t-)) .
\end{equation*}
If $U_{a \concat j}(t) < N$,
the new individual labeled $a \concat j$ is produced,
and $\kappa$ determines which label is associated with the new location,
so
\begin{align*}
    X_{a \concat j}(t)
    &=
    X_a(t-) + (1 - \kappa) y(X_a(t-), z) .
\end{align*}
On the other hand
if $U_{a \concat j}(t) \ge N$, then $X_a$ is unchanged and $X_{a \concat j}$ is undefined,
so
\begin{align} \label{eqn:X_line_of_descent}
    X_a(t)
    &=
    X_a(t-) + \kappa y(X_a(t-), z) \ind_{U_{a \concat j}(t) < N} .
\end{align}
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.
Combining these, $X_a$, for $\tau_a \le T < \sigma_a$, solves the equation 
\begin{align*}
    X_a(T)
    &=
    \int_{[\tau_a, T) \times [0, \infty) \times \IR \times [0, 1]}
    y(X_a(t-), z)
    \kappa
    \ind_{
        U_a(t) + v \ell(X_a(t-), y(X_a(t-), z), \eta_t)
    }
    d\Pi_a(s, v, z, \kappa) .
\end{align*}
